---
title: "DATA501 Assignement 2 - Corvin Idler - ID 300598312"
output:
  pdf_document: default
date: "2024-08-13"
---

This is a PDF generated from a RMarkdown file for Assignment 2 of the DATA 501 class 2024 from Victoria University Wellington
https://www.wgtn.ac.nz/courses/data/501/2024/offering?crn=33170

The repository underpinning this file and assignment can be found at https://github.com/econdatatech/distancemeasures/

I created an R package hosted at the above URL that will be loaded with the following lines of code

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,  warning = FALSE, message = FALSE)
library(devtools)
install_github("econdatatech/distancemeasures",force=TRUE)
```
The user will provide the program with a data set as well as a model (which is an object of class lm)
Three measures of influence will be calculated and plotted"

* Cooks Distance Measure (Cook, 1977) 
* DFFITS (Welsch and Kuh, 1977; Belsley, 1980) 
* Hadis Influence Measure (Hadi, 1992) 

To test the package we can use the data from the famous car package and fit a linear model and then plot some distance/influence measures

```{r packageuse, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(distancemeasures)
data(mtcars)
# Fit a linear regression model
# Predicting 'mpg' (miles per gallon) based on 
# 'disp' (displacement), 'hp' (horsepower), and 'wt' (weight)
model <- lm(mpg ~ disp + hp + wt, data = mtcars)
distances(mtcars,model)
```






