---
title: "DATA501 Assignement 2 - Corvin Idler - ID 300598312"
output:
  pdf_document: default
date: "2024-08-13"
bibliography: references.bib 

---
# Introduction

This is a PDF generated from a RMarkdown file for Assignment 2 of the DATA 501 class 2024 from Victoria University Wellington
https://www.wgtn.ac.nz/courses/data/501/2024/offering?crn=33170

The repository underpinning this file and assignment can be found at https://github.com/econdatatech/distancemeasures/

## Install instructions
I created an R package hosted at the above URL that will be loaded with the following lines of code

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,  warning = FALSE, message = FALSE)
suppressWarnings({ 
library(devtools)
install_github("econdatatech/distancemeasures",force=TRUE)
}) 
```
## Purpose and execution example
The user will provide the program with a data set as well as a model (which is an object of class lm)
Three measures of influence will be calculated and plotted"

* Cooks Distance Measure (Cook, 1977) 
* DFFITS (Welsch and Kuh, 1977; Belsley, 1980) 
* Hadis Influence Measure (Hadi, 1992) 

For sake of consistency and due to the discussion in @Hadi1992 (page 14) I decided to not include any cutoff values for the various influence measures. 

To test the package we can use the data from the famous car package and fit a linear model and then plot some distance/influence measures

```{r packageuse, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(distancemeasures)
data(mtcars)
# Fit a linear regression model
# Predicting 'mpg' (miles per gallon) based on 
# 'disp' (displacement), 'hp' (horsepower), and 'wt' (weight)
model <- lm(mpg ~ disp + hp + wt, data = mtcars)
distances(mtcars,model)
```

# Implementation details
The user facing function distances() makes use of the following helper functions (one for each distance/influence measure):

## Cooks distance
Below implementation is based on @Cook1977 (page 16ff.)

```{r cookscode, include=TRUE}
# based on https://doi.org/10.1080/00401706.1977.10489493
# and https://github.com/SurajGupta/r-source/blob/master/src/library/stats/R/lm.influence.R
cooks_distance_lm <- function(model) {
  # as per page 15 of Cook 1977 (above equation 1)
  resid <- stats::weighted.residuals(model) # to allow for weighted regression
  # residual degrees of freedom (number of observ.
  # minus number of regression coefficients)
  df <- stats::df.residual(model)
  sd <- sqrt(stats::deviance(model) / df)
  # diagonals of the ‘hat’ matrix.
  hat <- stats::lm.influence(model, do.coef = FALSE)$hat

  p <- model$rank
  # equation 7 in Cook 1977 page 16
  D <- ((resid / (sd * sqrt((1 - hat))))^2 * hat / (p * (1 - hat)))
  D[is.infinite(D)] <- NaN
  return(D)
}
```


## dffits
This implementation is based on @Belsley1980 and @Chatterjee2015
```{r dffitscode, include=TRUE}

# based on https://avys.omu.edu.tr/storage/app/public/rezzanu/141865/[David_A._Belsley,_Edwin_Kuh,_Roy_E._Welsch]_Regre(BookFi.org).pdf
# and https://github.com/SurajGupta/r-source/blob/master/src/library/stats/R/lm.influence.R
dffits_lm <- function(model) {
  hat <- stats::lm.influence(model, do.coef = FALSE)$hat
  sigma <- stats::lm.influence(model, do.coef = FALSE)$sigma
  res <- stats::weighted.residuals(model)
  # based on equation 2.11 on page 15 of Belsley 1980
  # re-written to avoid one sqrt(1-hat)
  dffits <- res * sqrt(hat) / (sigma * (1 - hat))
  dffits[is.infinite(dffits)] <- NaN
  return(dffits)
}
```

## Hadi
Ali Haid criticises in his 1992 paper that all prexisting influence measure only assessed the influence on a specific regression result, he on the other hand proposes "a measure of overall potential influence" @Hadi1992. The implementation below is based on the formula on page 113 in @Chatterjee2015. The formula tries to measure "outlyingness" and X-space (first term) as well as response variable space (second term) @Chatterjee2015.
```{r hadicode, include=TRUE}
#based on https://ideas.repec.org/a/eee/csdana/v14y1992i1p1-27.html
#and https://sadbhavnapublications.org/research-enrichment-material/2-Statistical-Books/Regression-Analysis-by-Example.pdf
hadi_lm <- function(model) {
  h <- stats::hatvalues(model)
  # based on sentence under equation 3.7 "normalized residuals". 
  di <- stats::residuals(model) / sqrt(sum(stats::residuals(model)^2))
  p <- length(stats::coef(model)) - 1
  result <-(h / (1 - h) + (p + 1) / (1 - h) * di^2 / (1 - di^2))
  result[is.infinite(result)] <- NaN
  return(result)
}
```
# Bibliograhy


